---
title: "INFO550_HW3"
author: "Limeng Wan"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, fig.width = 6, fig.height = 6
)
```

## 1. Data cleaning
```{r}
candat <- read.csv("./dataR2.csv")
n.missing <- sum(is.na(candat))
n.row <- nrow(candat)
candat$Classification <- factor(candat$Classification, labels = c("Controls", "Patients"))
freq <- as.data.frame(table(candat$Classification))
names(freq) = c('Classification', "Frequency")
```
There are total `r n.row` participants in the study with `r freq$Frequency[1]` healthy controls and `r freq$Frequency[2]` breast cancer patients. There are `r n.missing` missing values in the data set.


## 2. Summary statistics
```{r}
descriptive <- t(apply(candat[-10], 2, summary))
row.names(descriptive) <- colnames(candat)[-10]
t.row.name <- c("Age (years)", "BMI (kg/$m^2$)","Glucose (mg/dL)",
                 "Insulin ($\\mu$U/mL)", "HOMA","Leptin (ng/mL)", 
                 "Adiponectin ($\\mu$g/mL)","Resistin (ng/mL)","MCP-1(pg/dL)")
rownames(descriptive) <- t.row.name
knitr::kable(descriptive, digits = 4, caption = "Table 1. Summary Statistics", escape = FALSE)
```


## 3. Shapiro-Wilk test for normality for each continuous variable 
```{r}
SWtest <- apply(candat[-10],2,function(x) shapiro.test(x))
SWtest_sum = data.frame(Reduce(rbind, SWtest))
row.names(SWtest_sum) <- t.row.name
knitr::kable(SWtest_sum[c(-3,-4)], digits=4, col.names = c("W statistic", "p-value"), caption ="Table 2. Shapiro-Wilk Test Results")
```
<!-- The null-hypothesis of this test is that the population is normally distributed. Thus, if the p value is less than the chosen alpha level, then the null hypothesis is rejected and there is evidence that the data tested are not normally distributed. The chosen alpha level is 0.05 for all statistical tests. The data is normal if the p-value is above 0.05. -->
The p-values for the Shapiro-Wilk test for all continuous variables are less than 0.05, which indicates that all continuous variables do not follow normal distributions.


## 4. Mann-Whitney U tests for distribution differences between two populations
```{r}
# Mann-Whitney U tests
MWtest <- apply(candat[-10],2,function(x) wilcox.test(x~candat$Classification))
MWtest_sum = data.frame(Reduce(rbind, MWtest))
row.names(MWtest_sum) <- t.row.name
knitr::kable(MWtest_sum[c(-2,-4,-5,-6,-7)], col.names = c("U statistic", "p-value"), caption ="Table 3. Mann-Whitney U Test Results")
```
<!-- Under the null hypothesis H0, the distributions of both populations are equal. The alternative hypothesis H1 is that the distributions are not equal.-->
Since the continuous variables are not normal, use nonparametric Mann-Whitney U tests to test the distribution of two populations for each continuous variable. Based on the result table, Glucose, Insulin, HOMA and Resistin differs significantly between two populations because p-values are less than 0.05.


## 5. UVA Logistic Regression Model
```{r}
library(sjPlot)
library(sjmisc)
library(sjlabelled)
# Univariate analysis
log_res <- c()
log_res <- apply(candat[-10],2,function(Var) glm(Classification ~ Var, data=candat, family=binomial(link="logit")))
tab_model(log_res, dv.labels = colnames(candat)[-10], show.ci = F)
```
Based on the univariate logistic regression analysis, Glucose, Insulin, HOMA and Resistin have significant relationships with the odds of being breast cancer patients. 


## 6. MVA Logistic Regression Model
```{r}
# Multivariate analysis
log_1 <- glm(Classification ~ Age + BMI + Glucose + Insulin + HOMA + Resistin + MCP.1, data=candat, family=binomial(link="logit"))
log_2 <- glm(Classification ~ BMI + Glucose + Insulin + HOMA + Resistin + MCP.1, data=candat, family=binomial(link="logit"))
log_3 <- glm(Classification ~ BMI + Glucose + Insulin + HOMA + Resistin, data=candat, family=binomial(link="logit"))
log_4 <- glm(Classification ~ BMI + Glucose + Insulin + Resistin, data=candat, family=binomial(link="logit"))
log_5 <- glm(Classification ~ BMI + Glucose + Resistin, data=candat, family=binomial(link="logit"))
tab_model(log_1,log_2,log_3,log_4,log_5,dv.labels = c("MVA Model 1","MVA Model 2","MVA Model 3","MVA Model 4","MVA Model 5"), show.ci = F)
```
Base on the multivariate logistic regression model, BMI, Glucose and Resistin have significant relationships with the odds of being breast cancer patients. 


## 7. Working directory and session info
```{r}
getwd()
devtools::session_info()
```
